Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-22 methods/subsections/machine learning.tex
Updates from Overleaf

\section{Artificial Neural Networks (ANNs)}
Artificial Neural Networks (ANNs) are computational models inspired by biological neural systems. They use interconnected layers of artificial neurons to process information and perform tasks like pattern recognition, classification, and regression. ANNs adjust the connections, or "weights," between neurons during learning to improve performance. The concept of neural networks emerged in the mid-20th century with the work of Warren McCulloch and Walter Pitts. Still, modern neural networks began with Frank Rosenblatt's perceptron in 1957, a single-layer network able to learn linearly separable patterns.

>>>>

\section{Machine Learning Modeling}

This section outlines the methodical procedure of creating, training and assessing algorithms that enable data-driven learning and decision-making. The process involves algorithm selection, model architecture design, data preprocessing and transformation, and feature extraction. The objective is to minimize prediction error by iteratively fine-tuning the model's parameters. The selection between supervised, unsupervised, or reinforcement learning techniques depends on the problem's specifics and data structure. It is crucial to evaluate, refine, and deploy the model to ensure its effective performance with new data in real-world applications.

I advocate for exposure as the primary method of learning music, highlighting the importance of various musical experiences for a comprehensive understanding of the art form. Engagement with diverse musical styles, genres, techniques, and learning methods fosters well-rounded musical proficiency, encouraging creativity and experimentation.

Considering the task at hand, we suggest using a Triple Siamese Network, an architecture proven efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The goal is to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}.

Using the proposed triple siamese network and online triplet mining in a self-supervised setting, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

At a superficial level, our approach could be described as cognitive modeling, given that it seeks to approximate human problem-solving and mental processing through a computerized model.

\subsection{Artificial Neural Networks (ANNs)}
Artificial Neural Networks (ANNs) are computational models inspired by biological neural systems. They use interconnected layers of artificial neurons to process information and perform tasks like pattern recognition, classification, and regression. ANNs adjust the connections, or "weights," between neurons during learning to improve performance. The concept of neural networks emerged in the mid-20th century with the work of Warren McCulloch and Walter Pitts. Still, modern neural networks began with Frank Rosenblatt's perceptron in 1957, a single-layer network able to learn linearly separable patterns.

### 2023-05-22 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\section{Machine Learning Modeling}

This section outlines the methodical procedure of creating, training and assessing algorithms that enable data-driven learning and decision-making. The process involves algorithm selection, model architecture design, data preprocessing and transformation, and feature extraction. The objective is to minimize prediction error by iteratively fine-tuning the model's parameters. The selection between supervised, unsupervised, or reinforcement learning techniques depends on the problem's specifics and data structure. Evaluating, refining, and deploying the model to ensure its effective performance with new data in real-world applications is crucial.

I advocate for exposure as the primary method of learning music, highlighting the importance of various musical experiences for a comprehensive understanding of the art form. Engagement with diverse musical styles, genres, techniques, and learning methods fosters well-rounded musical proficiency, encouraging creativity and experimentation.

Considering the task at hand, we suggest using a Triple Siamese Network, an architecture proven efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The goal is to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}.

Using the proposed triple siamese network, and online triplet mining in a self-supervised setting, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

@@

\subsection{Siamese Networks}

A Siamese Network, a deep learning architecture introduced by Bromley and LeCun in the early 1990s \cite{Bromley1993SignatureNetwork}, is designed explicitly for tasks involving similarity or comparison between two input instances. Its name derives from the structure, which consists of two or more identical subnetworks connected in parallel and joined at the output layer. These subnetworks share the same architecture, weights, and hyperparameters, allowing for more efficient memory usage and computational complexity.

This design is particularly effective for learning from limited or imbalanced datasets, as it focuses on learning the similarity metric rather than specific features of individual classes.

Each subnetwork processes an input independently, and its outputs are combined and further processed to yield a single similarity score. The shared weights during training enable the model to learn an invariant representation for the input instances, improving its efficiency in comparing and contrasting them. Siamese Networks employ specialized loss functions, such as contrastive or triplet loss, which minimize the distance between similar input pairs and maximize the distance between dissimilar ones. This approach encourages the network to learn a meaningful similarity metric.

\subsubsection{Triplet Siamese Networks}

Triplet Siamese Networks, an extension of the Siamese Network architecture, involve comparing three input instances instead of two, aiming to learn an embedding space where similar instances are close and dissimilar instances are distant. The input consists of an anchor, a positive instance (same class as the anchor), and a negative instance (different class). An identical subnetwork with shared architecture, weights, and hyperparameters processes each instance.

The triplet loss function guides learning, minimizing the distance between the anchor and positive instances while maximizing the distance between the anchor and negative instances. A margin parameter in the loss function ensures a minimum separation between positive and negative instances in the embedding space.

Due to shared weights, Triplet Siamese Networks offer efficient memory usage and computational complexity.

@@

In this equation, triplet loss function with a variable margin:

>>>>

@@

\subsection{Siamese and Triplet Siamese Networks}

A Siamese Network is a deep learning architecture introduced by Bromley and LeCun in the early 1990s \cite{Bromley1993SignatureNetwork}. This architecture is explicitly designed for tasks requiring comparing or assessing similarity between two input instances. Its name is derived from the network structure, which consists of two or more identical subnetworks connected in parallel and joined at the output layer. These subnetworks share the same architecture, weights, and hyperparameters, allowing more efficient memory usage and computational complexity.

Siamese Networks are especially effective for learning from limited or imbalanced datasets, as they focus on learning a similarity metric rather than the specific features of individual classes. Each subnetwork processes an input independently, and its outputs are then combined and further processed to yield a single similarity score. The shared weights during training enable the model to learn an invariant representation for the input instances, thereby enhancing its efficiency in comparing and contrasting them. To achieve this, Siamese Networks employ specialized loss functions, such as contrastive or triplet loss, which aim to minimize the distance between similar input pairs and maximize the distance between dissimilar ones.

An extension of this architecture is the Triplet Siamese Network, which involves comparing three input instances instead of two. This network aims to learn an embedding space in which similar instances are close to each other and dissimilar instances are distant. The input to this network consists of an anchor, a positive instance (from the same class as the anchor), and a negative instance (from a different class). Each instance is processed by an identical subnetwork that shares its architecture, weights, and hyperparameters with the others.

The triplet loss function used in this network guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space. Due to the shared weights among the subnetworks, Triplet Siamese Networks also offer efficient memory usage and computational complexity.

@@

In this equation, the triplet loss function with a variable margin:
